---
title: "NYTViewer"
author: "QiHan Zhao"
date: '2022-09-05'
output:
  pdf_document: default
  html_document: default
runtime: shiny
---

```{r librarySetup}
library(tidyverse)
library(httr2)
library(jsonlite)
library(glue)
```

### Task 1 - Figuring out the NY Times Article Search API

```
https://api.nytimes.com/svc/search/v2/articlesearch.json?
```

* `begin_date=20190905`

* `end_date=20190905`

* `fq=document_type:article+AND+print_page:1+AND+print_section:A`

* `page=0`

<br/>    

```{r api_task1}
api.key <- "zNCIxN6GzTKA4UZ2rhuaktGqMuYmgwOX"
api <- "https://api.nytimes.com/svc/search/v2/articlesearch.json?"
date <- "20190905"
pageNum <- 0
test_json <- jsonlite::read_json(
  glue::glue(
    "https://api.nytimes.com/svc/search/v2/articlesearch.json?fq=document_type:article+AND+print_page:1+AND+print_section:A&begin_date={date}&end_date={date}&page={pageNum}&api-key={api.key}"
  )
)
```


### Task 2 - Getting data from the NY Times Article Search API
For this task, I started off writing the abstractions: laying out the important
event handlers for the main function get_nyt_articles(), and then writing out the 
code for each function (they are also in sequential order that is aligned with the 
order I wrote them). Because we need to first check whether the passed-in month, 
day arguments are in valid range, I wrote a simple boolean function to do that
for me; I did not use the stopifnot() function because I thought the way I titled
my function is more informative and readable in the context of a if statement.
Then, I wrote a function that formats year, month, day into usable string format
for the API argument begin_date & end_date; I tried to use a different API argument, 
which would not require me to format the dates but using it returns unwanted headlines; 
I suspected that it was because that argument also required a time stamp value so 
I sticked with using the begin_date & end_date. After that, I wrote a function that
handles the requesting and downloading matched json file. I used response\$meta\$hits
to check how many total hits there are and whether the function can terminate after 
current page download; at the end of the repeat loop, I also called Sys.sleep(6) 
to space out the API calls to comply with the requirement of not sending requests
too frequently. My last helper function is a data wrangling function that helps 
me to clean the downloaded json files into a dataframe with relevant columns, per
the github instruction, such as byline, headline, and webURL etc. In the end, I 
called each of the helper function inside the get_nyt_articles() function. The
function first checks whether month, day are in the valid range; if they are not, 
an informative message is printed in the console. Then the function proceeded to 
make API requests and download the json files. If the request has no hits, the 
function returns NULL and a cleaned dataframe of the downloaded file otherwise. 
```{r}
# @param: checked argument, lower bound, upper bound
# @return: whether the argument is in valid range
arg_not_between <- function(val, low, high){
  if ((low <= val && val <= high)) {
    return (FALSE)
  }
  return (TRUE)
}
# @param: year, month, day
# @return: date in string format usable for the API call
#         for example: 2019/2/15 -> 20190215
format_date_entry <- function(year, month, day) {
  if (month < 10) {
    MONTH <- paste0("0", month)
  } else {
    MONTH <- month
  }
  if (day < 10) {
    DAY <- paste0("0", day)
  } else {
    DAY <- day
  }
    
  return (paste0(year, MONTH, DAY))
}
# @param: date, API key
# @return: matched NYT articles in json file (nested list)
nyt_api_req <- function(date, api_key) {
  tbl <- list()
  page <- 0
  # initialize the number of hits left to an arbitrary number
  documentsLeft <- 10
  repeat {
    cat("Downloading Page ", page, "\n")
    resp <- jsonlite::read_json(
      glue::glue("https://api.nytimes.com/svc/search/v2/articlesearch.json?fq=document_type:article+AND+print_page:1+AND+print_section:A&begin_date={date}&end_date={date}&page={page}&api-key={api_key}")
    )
    if (page == 0) {
      documentsLeft = resp$response$meta$hits
    }
    # check how many documents are left
    documentsLeft = documentsLeft - length(resp$response$docs)
    # update the users on the download status
    cat("Articles downloaded:  ", length(resp$response$docs), "\n")
    cat("Articles left:  ", documentsLeft, "\n")
    tbl <- c(tbl, resp)
    
    if (documentsLeft <= 0) {
      break
    }
    page <- page + 1
    Sys.sleep(6)
  }
  tbl
}
wrangling_nyt_json <- function(df) {
  return (df %>%
    tibble() %>%
    tidyr::unnest_wider(., .) %>%
    select(docs) %>%
    filter(
      !purrr::map_lgl(docs, is.null)
    ) %>%
    tidyr::unnest_longer(docs) %>%
    tidyr::hoist(
      docs,
      webURL = "web_url",
      headline = c("headline", "main"),
      lead_paragraph = "lead_paragraph",
      byline = c("byline", "original"),
      source = "source"
    ))
}
get_nyt_articles = function(year, month, day, 
                            api_key = "zNCIxN6GzTKA4UZ2rhuaktGqMuYmgwOX") {
  if(arg_not_between(as.integer(month), 1, 12)) stop("Month is not in valid range!")
  if(arg_not_between(as.integer(day), 1, 31)) stop("Day is not in valid range!")
  
  DATE <- format_date_entry(as.integer(year), as.integer(month), as.integer(day))
  
  APIdata_in_json <- nyt_api_req(DATE, api_key)
  if(APIdata_in_json$response$meta$hits == 0) {
    return (NULL) 
  }
  
  final.df <- wrangling_nyt_json(APIdata_in_json)
  final.df
}
```

<br/>

### Task 3 - Shiny Front End
In task3, I used text input box UI for users to enter their desired date, which 
is then converted to integers before passed into the API request functions. The
default values for the dates is my birthday when I entered college, and I also
added a Request action button that would trigger the API request when the user
clicked it. Once the button is clicked, the get_nyt_articles() function from 
Task 2 is called to download matched headlines; I wrapped the download function
into withProgress(), which would update the users on what the App is doing. If
no headline exists on thatdate the App will display a message box that prompts 
the user to try a different date. To format the matched headlines, I wanted to 
turn each headline into an actionlink, embedded in a rendered dataframe for 
aeshtetic purpose, that would trigger a modal dialog box upon being clicked. In 
order to achieve this, I searched up ways to inporcorate JavaScript lines to 
my shiny functions because I thought this functionality might be beyond the ability
of Shiny and thus entailed the use of JavaScript. Eventually I found a code on
StackOverflow that achieves what I planned to do(https://stackoverflow.com/questions/45739303/r-shiny-handle-action-buttons-in-data-table). 
This approach is convenient because, to some extent, it saves me from the process
of creating observers for each headline since all of the headline can then
be accessed through one "input id"--that is, input$select_headline--, and yet
each actionlink still has its headline as its unique id that can be used to 
filter for matched dataframe row to create popup modal dialog box. After I implemented
this feature, I wrote a observer that creates a popup modal dialog box based on 
which headline the users clicked; I passed in some CSS values to center the action
button that would take the user to the actual article in the modal dialog box. 


```{r}
library(shiny)
shinyApp(
  ui = fluidPage(
    titlePanel("NYTimes API"),
    sidebarLayout(
      sidebarPanel(
        h4("Request For Publication on this date"),
        textInput("year", "Year", value = 2019),
        textInput("month", "Month", value = 9),
        textInput("day", "Day", value = 5),
        h4("Enter Your API key (if applicable)"),
        textInput("APIKEY", ""),
        actionButton("requestData", "Request")
      ),
      mainPanel(
        DT::dataTableOutput("headlineTbl")
      )
    )
  ),
  server = function(input, output, session) {
    
    HeadlineToLink <- function(headlines){
      
      # this function turned into the action link UI into strings
      # which will then be rendered as vanilla actionlink when the App runs. 
      # Passed in the javascript function at the end, so that I can access which 
      # actionlink is pressed through input$select_headline, which returns the id
      # of each actionlink (which is set to be its headline)
      # CODE REFERENCE: https://stackoverflow.com/questions/45739303/r-shiny-handle-action-buttons-in-data-table
      HL.list <- character(length(headlines))
      for (i in seq_len(length(headlines))) {
        HL.list[i] <- as.character(actionLink(paste0(headlines[i]), 
                                              paste0(headlines[i]),
                                              onClick = 'Shiny.onInputChange(\"select_headline\",  this.id)'))
      }
      HL.list
    }
    state = reactiveValues(
      observers = list()
    )
    
    request.data = reactiveValues(tbl = NULL)
    
    download_api_data <- function(){
      if(input$APIKEY != "") {
        request.data$tbl <- get_nyt_articles(input$year,
                                             input$month,
                                             input$day,
                                             api_key = input$APIKEY)
      } else {
        request.data$tbl <- get_nyt_articles(input$year,
                                             input$month,
                                             input$day)
      }
    }
    # @return: dataframe with infomation on byline, title, paragraph of 
    #          selected headline
    extractDataRow <- function(hl) {
      req(request.data$tbl)
      
      request.data$tbl %>% 
        filter(headline == hl)
    }
    observeEvent(input$requestData, {
      
      # Progress bar that updates the user on the current process of the function
      withProgress(message = "Loading", value = 0, {
        incProgress(0.3, detail = "Requesting & Downloading API data")
        download_api_data()
        if (is.null(request.data$tbl)) {
          incProgress(0.7, detail = "")
          showNotification("It seems that your request has no hits, Please change a date",
                           type = "message",
                           duration = 2)
        } else{
          incProgress(0.5, detail = "Rendering Headline Table")
          request.data$headline <- data.frame(
            date = paste0(input$year, "/", input$month, "/", input$day),
            headlines = HeadlineToLink(request.data$tbl$headline)
            )
          Sys.sleep(1)
        }
      })
      
      # set escape to FALSE so that the actionLink UI string is rendered instead
      # of treating as actual string contents
      output$headlineTbl <- DT::renderDataTable(
        request.data$headline, server = FALSE, escape = FALSE, selection = 'none'
        )
    })
    
    # create the modal dialog box based on the headline that the user clicked
    # input$select_headline gives the headline that they click and is used to
    # matched the correct row inside the NYT articles dataframe, which is then used
    # to create the modal dialog box. 
    # at the end, use CSS to center the action button
    observeEvent(input$select_headline, {
      request.data$selectedHL <- extractDataRow(input$select_headline)
      showModal(modalDialog(
        h4("Title"),
        p(request.data$selectedHL$headline),
        h4("Byline"),
        em(request.data$selectedHL$byline),
        h4("A peek on the content"),
        p(request.data$selectedHL$lead_paragraph),
        div(actionButton("ToSource", "I want to read it",
                     onClick = paste0("window.open('", 
                                      request.data$selectedHL$webURL,
                                      "')")), style = "display: flex; align-items: center; justify-content:center;")
      ))
    })
  }
)
```